---
title: "Scratch"
output: html_document
---

```{r}
library(ggplot2)
library(dplyr)
library(stringr)
library(tidyr)
```


```{r}
cfb <- read.csv("http://www.stat.cmu.edu/cmsac/sure/materials/data/regression_projects/cfb_2019_games.csv")
head(cfb)
```

```{r}
cfb

unique(cfb$home_conference)
```

Algamation of all data transformations

```{r}
cfb$home_1sthalf <- cfb$home_1_pts + cfb$home_2_pts 
cfb$home_2ndhalf <- cfb$home_3_pts + cfb$home_4_pts

cfb$away_1sthalf <- cfb$away_1_pts + cfb$away_2_pts
cfb$away_2ndhalf <- cfb$away_3_pts + cfb$away_4_pts

cfb$margin_1sthalf <- cfb$home_1sthalf - cfb$away_1sthalf
cfb$margin_game <- cfb$home_points - cfb$away_points

cfb$comeback <- (cfb$margin_game / cfb$margin_1sthalf) < 0

cfb$overtime <- !((cfb$home_1_pts + cfb$home_2_pts + cfb$home_3_pts + cfb$home_4_pts) == cfb$home_points |
                    (cfb$away_1_pts + cfb$away_2_pts + cfb$away_3_pts + cfb$away_4_pts) == cfb$away_points)
```

Some conference data is not included in this list, such as the "Big Sky", which gave us NA values for their away conference when they played teams in this dataset. So for future reasons, we will change all teams without a conference in the "Other" category.

```{r}
for (i in 1:length(cfb$X)) {
  if (is.na(cfb$away_conference[i])) {
    cfb$away_conference[i] = "Other"
  }
}

```



All games have an excitement level besides these two games:
```{r}
cfb[is.na(cfb$excitement_index),]

cfb_ei <- cfb[-c(268,625),]
```



Thought about looking at time but when we realized they were all in GMT, the effort to find the timezone for each stadium/home team and then change it is something we did not have time before.

```{r}
head(
  cfb %>%
  separate(col = start_date, into = c("date", "time"), sep = "T") %>%
  separate(col = date, into = c("year", "month", "day"), sep = "-") %>%
  separate(col = time, into = c("hour", "minutes"), sep = ":")
)
```


```{r}
ggplot(cfb_ei, aes(x = excitement_index)) +
  geom_histogram(binwidth = 1, color = "black", fill = "gold4") +
  stat_function(fun = dnorm) +
  labs(title = "Normality of Excitement Index Variable",
       x = "Excitement Index") +
  theme_bw()

ggplot(cfb_ei, aes(x = excitement_index)) +
  geom_histogram(binwidth = .5)

ggplot(cfb_ei, aes(x = excitement_index)) +
  stat_ecdf()
```

```{r}
ggplot(cfb_ei, aes(sample = excitement_index)) + 
  stat_qq(color = "gold4") + 
  stat_qq_line(color = "black") +
  labs(title = "Q-Q Plot of Excitement Index")
```






```{r}
ggplot(cfb_ei, aes(excitement_index, score_difference)) +
  geom_point(aes(color = comeback, shape = overtime, size = overtime), alpha = 0.7) +
  geom_smooth(method = "lm") +
  scale_shape_manual(values = c(16, 3)) +
  ylim(0, 80)

ggplot(cfb_ei, aes(excitement_index, score_difference)) +
  geom_point(aes(color = comeback, shape = overtime, size = overtime), alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylim(0, 80)

ggplot(cfb_ei[cfb_ei$conference_game == FALSE, ], aes(excitement_index, score_difference)) +
  geom_point(aes(color = comeback, shape = overtime, size = overtime)) +
  geom_smooth(method = "lm") +
  ylim(0, 80)

ggplot(cfb_ei[cfb_ei$conference_game == FALSE, ], aes(excitement_index, score_difference)) +
  geom_point(aes(color = comeback, shape = overtime, size = overtime)) +
  geom_smooth(method = "lm") +
  scale_shape_manual(values = c(16, 3)) +
  ylim(0, 80)

ggplot(cfb_ei[cfb_ei$conference_game == TRUE, ], aes(score_difference, excitement_index)) +
  geom_point(aes(color = comeback, shape = overtime, size = overtime), alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylim(0, 10) +
  labs(title = "Competitive Games Breed the Highest Excitement",
       subtitle = "Close Games, Overtimes, and Comebacks",
       x = "Score Difference", 
       y = "Excitement Index",
       color = "Comeback") +
  theme_bw()
ggsave("FinalTakeaway_ScoreDifferencevsEI.jpeg", width = 8, height = 6)


```







```{r}
ggplot(cfb_ei, aes(home_points+away_points, excitement_index)) +
  geom_point(color = "gold4") +
  geom_smooth() +
  xlab("Total Points") +
  ylab("Excitement Index") +
  ggtitle("Home+Away Points, Smoothed") +
  theme_bw()
ggsave("Points_vs_EI.jpg", width = 10, height = 6) 
ggplot(cfb_ei, aes(home_4_pts+away_4_pts, excitement_index)) +
  geom_point(color = "gold4") +
  geom_smooth() +
  xlab("Total Points") +
  ylab("Excitement Index") +
  ggtitle("4th Quarter Total Points, Smoothed") + 
  theme_bw()
ggsave("4thQPoints_vs_EI.jpg", width = 10, height = 6)

  
```






```{r}
cfb_ei %>%
  group_by(week) %>%
  summarize(avg_excitement = mean(excitement_index), se = sd(excitement_index) / sqrt(n())) %>%
  ggplot(aes(x = week, y = avg_excitement)) +
  geom_col(color = "black", fill = "gold4") +
  labs(x = "Week", 
       y = "Total Excitement Level",
       subtitle = "95% Confidence Intervals") +
  geom_errorbar(aes(ymin=avg_excitement - (1.96*se), ymax=avg_excitement + (1.96*se)), width=.2,
                 position=position_dodge(.9)) +
  theme_bw()
ggsave("AvgEIbyWeek.jpg", width = 10, height = 6)


cfb_ei %>%
  group_by(week) %>%
  summarize(avg_excitement = sum(excitement_index)) %>%
  ggplot() +
  geom_col(aes(x = week, y = avg_excitement), color = "black", fill = "gold4") +
  labs(x = "Week", 
       y = "Total Excitement Level") +
  theme_bw()

```



```{r}
cfb_ei %>%
  group_by(comeback) %>%
  summarise(avg_ei = mean(excitement_index)) %>%
  ggplot(aes(comeback, avg_ei, fill = comeback)) +
  geom_col()

```

```{r}
cfb_ei %>%
  group_by(comeback) %>%
  summarise(avg_ei = mean(excitement_index), se = sd(excitement_index) / sqrt(n())) %>%
  ggplot(aes(comeback, avg_ei, fill = comeback)) +
  geom_col() +
  labs(x = "Comeback (After 1st Half)",
       y = "Average Excitement Index") +
  scale_fill_manual(values = c("darkgray", "gold4")) +
  geom_text(aes(label=round(avg_ei,2)), vjust=1.6, color="white", size=10) +
  theme(legend.position = "none") + 
  ggtitle("Games with Comebacks Garner High Excitement") +
  labs(subtitle = "95% Confidence Intervals") +
  geom_errorbar(aes(ymin=avg_ei - (1.96*se), ymax=avg_ei + (1.96*se)), width=.5,
                 position=position_dodge(.9)) +
  theme_bw()

  

cfb_ei %>%
  group_by(conference_game) %>%
  summarise(avg_ei = mean(excitement_index)) %>%
  ggplot(aes(conference_game, avg_ei, fill = conference_game)) +
  geom_col()
  
```

```{r}
conference_int <- cfb_ei %>%
  group_by(home_conference, away_conference) %>%
  summarise(num_games = n(),
            avg_ei = mean(excitement_index))


conference_int$combined = 0


for (i in 1:length(conference_int$num_games)) {
  for (j in 1:length(conference_int$num_games)) {
    if ((conference_int$home_conference[i] == conference_int$away_conference[j]) &&
        (conference_int$home_conference[j] == conference_int$away_conference[i])) {
      conference_int$combined[i] = ((conference_int$avg_ei[i] * conference_int$num_games[i]) + 
                                      (conference_int$avg_ei[j] * conference_int$num_games[j])) / 
                                      (conference_int$num_games[i] + conference_int$num_games[j])
    }
  }
}

conference_int[order(-conference_int$combined),]

conference_int$title = paste(conference_int$away_conference, "at", conference_int$home_conference)


exciting_conf_matchups = conference_int[conference_int$combined > 5,]
exciting_conf_matchups$home_conference[1] = "AA"
exciting_conf_matchups$away_conference[5] = "AA"
exciting_conf_matchups$home_conference[5] = "MW"
exciting_conf_matchups$home_conference[6] = "MW"
exciting_conf_matchups$away_conference[1] = "MW"
exciting_conf_matchups$away_conference[3] = "MW"

exciting_conf_matchups$title = paste(exciting_conf_matchups$away_conference, 
                                     "at", 
                                     exciting_conf_matchups$home_conference)



ggplot(data = exciting_conf_matchups[order(-exciting_conf_matchups$combined),]) +
  geom_col(aes(title, avg_ei)) +
  theme(axis.text.x=element_text(angle=20,hjust=1))

```


```{r}
test <- cfb %>%
  group_by(venue) %>%
  summarise(num_games = n(),
            avg_ei = mean(excitement_index))

test[order(-test$avg_ei) ,]

test <- cfb %>%
  group_by(home_team) %>%
  summarise(num_games = n(),
            avg_ei = mean(excitement_index))

head(test[order(-test$avg_ei) ,])
```

```{r}
mean(cfb_ei[cfb_ei$margin_game > 0,]$excitement_index)
mean(cfb_ei[cfb_ei$margin_game < 0,]$excitement_index)

mean(cfb_ei[cfb_ei$margin_game > 0 & cfb_ei$margin_game < 30,]$excitement_index)
mean(cfb_ei[cfb_ei$margin_game < 0 & cfb_ei$margin_game > -30,]$excitement_index)

```





```{r}
cfb_ei %>%
  mutate(home_win = cfb_ei$margin_game > 0) %>%
  select(excitement_index, home_win) %>%
  group_by(home_win) %>%
  summarise(avg_ei = mean(excitement_index)) %>%
  ggplot(aes(y=avg_ei, x=home_win)) + 
  geom_col() +
  geom_text(aes(label=round(avg_ei,2)), vjust=1.6, color="white",
            position = position_dodge(0.9), size=10) +
  xlab("Home Team Win") +
  ylab("Excitement Index") +
  theme_bw()

cfb_ei %>%
  mutate(home_win = cfb_ei$margin_game > 0) %>%
  select(excitement_index, comeback, home_win) %>%
  group_by(home_win, comeback) %>%
  summarise(avg_ei = mean(excitement_index), se = sd(excitement_index) / sqrt(n())) %>%
  ggplot(aes(fill=comeback, y=avg_ei, x=home_win)) + 
  geom_col(position="dodge", stat="identity") +
  geom_text(aes(label=round(avg_ei,2)), vjust=3, color="white",
            position = position_dodge(0.9), size=3.5) +
  scale_fill_manual(values=c('gold4','darkgray')) +
  xlab("Home Team Win") +
  ylab("Excitement Index") +
  labs(title = "Trends of Home vs Away Wins Reversed when considering Comeback games",
       subtitle = "95% Confidence Intervals") +
  labs(fill = "Comeback") +
  geom_errorbar(aes(ymin=avg_ei - (1.96*se), ymax=avg_ei + (1.96*se)), width=0.5,
                 position=position_dodge(.9)) +
  theme_bw()

```

```{r}
head(cfb_ei[order(-cfb_ei$excitement_index),])
```

```{r}
cfb_ei[379,]
cfb_ei[471,]
cfb_ei[288,]

```


